---
title: "Week 10"
subtitle: "Factors"
author: "Michael W. Kearney📊<br/>School of Journalism <br/>Informatics Institute <br/>University of Missouri"
date: "<table style=\"border-style:none;padding-top:30px;\" class=\".table\">
  <tr>
    <th style=\"padding-right:75px!important\"><a href=\"https://twitter.com/kearneymw\"> <i class=\"fa fa-twitter\"></i> </a></th>
    <th style=\"padding-left:75px!important\"><a href=\"https://github.com/mkearney\"> <i class=\"fa fa-github\"></i> </a></th>
  </tr>
  <tr style=\"background-color:#fff\">
    <th style=\"padding-right:75px!important\"><a href=\"https://github.com/mkearney\"> @kearneymw </a></th>
    <th style=\"padding-left:75px!important\"><a href=\"https://github.com/mkearney\"> @mkearney </a></th>
  </tr></table>"
output:
  xaringan::moon_reader:
    css: ['robot', 'robot-fonts']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, collapse = TRUE, comment = ">")
options(htmltools.dir.version = FALSE)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
head <- function(x, n = 6) {
  x <- utils::head(x, n)
  knitr::kable(x, format = "html")
}
```

# \#rstats

Visualizations and data wrangling done using the [tidyverse](https://tidyverse.org).

+ **{ggplot2}**: data visualizations
+ **{dplyr}**: data wrangling

```{r}
## load the tidyverse
library(tidyverse)
```

---

# Factor

A **factor** is a psychometric term for variable. 

In statistics, factors describe the **latent** variables which we are attempting to measure. 


---

# Items

An **item** refers to a single question/prompt/response-provoking stimulus in a questionnaire.

In the context of a study, a **variable** refers to a construct (or factor) to be examined.

Variables can consist of one or more items from a questionnaire.


---

# Multi-item variables

When a variable (or factor) is measured using multiple items, we still want to represent it with one number.

**How can we represent a variable measured with 5 likert-type items using one number per observation (respondent)?**


---

# Example

Our variable of interest is extraversion/introversion. We measure it using four likert-type items:

+ I like talking to lots of people
+ I like attending events where I meet knew people
+ I am comfortable at a party where I don't know anyone else
+ Meeting my friends' friends makes me nervous


---

# Example responses

```{r}
## first person's responses
person1 <- c(1, 2, 1, 7)

## second person's responses
person2 <- c(6, 6, 5, 2)
```


---

# Factor analysis

To make sure we have **reliable** measures, we use factor analysis.

**Factor analysis** essentially finds the correlation between responses for similar items. 

There are lots of details and variations, but knowing this much will help you in the future!

