---
title: "Week 8"
subtitle: "Exam I Review"
date: "`r Sys.Date()`"
author: "Michael W. Kearney, PhD"
institute: |
  | School of Journalism
  | Informatics Institute
  | University of Missouri
output: 
  uslides::uslides_html:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
Sys.setenv(
  USLIDES_BOTTOMLEFT = "uslidelogos/bl.jpg",
  USLIDES_BOTTOMRIGHT = "uslidelogos/br.jpg",
  USLIDES_BANNER = "uslidelogos/bn.jpg",
  USLIDES_BACKGROUND = "uslidelogos/bg.jpg"
)
```

## Exam 1 Prep
+ Exam I
    - Identify level of measurement (nominal, ordinal, interval, ratio)
    - Identify appropriate R class (character/factor, integer, numeric, data.frame)
    - Interpret R code/output
    - Predict R output
    - Calculate [very simple] descriptive statistics
    - Select calculated descriptive statistics

# Levels of measurement

## Level of measurement
> + A likert-like item on a scale from 1 (strongly disagree) to 7 (strongly agree)
>     + **interval**
> + The weight of a person (in lbs.)
>     + **ratio**
> + The order in which horses finished the race
>     + **ordinal**

## Level of measurement
> + The name of a newspaper
>     + **nominal**
> + The temperature (in F)
>     + **interval**
> + The number of times an article says "fake news"
>     + **ratio**

## Level of measurement
> + Whether a movie is "rotten" or "fresh"
>     + **nominal**
> + What languages a journalist speaks
>     + **nominal**
> + How happy someone is on a scale from 1 (not happy) to 10 (very happy)
>     + **interval**


## Level of measurement
- **interval** A likert-like item on a scale from 1 (strongly disagree) to 7 (strongly agree)
- **ratio** The weight of a person (in lbs.)
- **ordinal** The order in which horses finished the race
- **nominal** The name of a newspaper
- **interval** The temperature (in F)

# R code/output

## R output
Predict the output: what is `y` equal to?
```{r, eval=FALSE}
x <- c(1, 2)
y <- x - 1
y
```
    a. `0`
    b. `c(0, 1)`
    c. `1`
    d. `c(1, 0)`

> + **b**

## R output
Predict the output: what is `y` equal to?
```{r, eval=FALSE}
x <- c(1, 2, -5, 2, 0, 14, NA)
y <- mean(x)
y
```
    a. `1.5`
    b. `2.33`
    c. `NA`
    d. `c(1, 2, -5, 2, 0, 14, NA)`

> + **c**

# R classes

## R class
Predict the output: what class is `x`?
```{r, eval=FALSE}
x <- c(1L, 2L, 3L)
class(x)
```
> + **integer**

## R class
Predict the output: what class is `x`?
```{r, eval=FALSE}
x <- c(0.25, -5.26, 2000)
class(x)
```
> + **numeric**

## R class
Predict the output: what class is `x`?
```{r, eval=FALSE}
x <- c("1", "2", "3")
class(x)
```
> + **character**

## R class
Predict the output: what class is `x`?
```{r, eval=FALSE}
x <- data.frame(
  x = c("a", "b", "c"),
  y = c("x", "y", "z")
)
class(x)
```
> + **data.frame**


# Descriptives

## Descriptive statistic
Predict the output: What is the standard deviation of x?
```{r, eval=FALSE}
x <- c(-50, -40, -30, -20, -10, 0, 10, 20, 30, 40, 50)
sd(x)
```
    a. 300
    b. -33.17
    c. 50
    d. 33.17

> + **d**

## Descriptive statistic
Predict the output: What is the mode of x?
```{r, eval=FALSE}
x <- c("a", "a", "b", "c", "c", "c")
sort(table(x), decreasing = TRUE)
```
    a. a
    b. b
    c. c

> + **c**

## Descriptive statistic
Predict the output: What is the median of x?
```{r, eval=FALSE}
x <- c(-50, 0, 1, 2, 3)
median(x)
```
    a. -50
    b. -8.8
    c. 0
    d. 1

> + **d**


