---
title: "Homework #2"
subtitle: "JOURN 8006"
author: "\\<enter your name here>"
date: "3/06/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

For this homework assignment, create a new folder and save the `2-descriptives.Rmd` and `2-data.csv` files in it.

## Data

Read in the data file (`2-data.csv`) using the code below.

```{r}
## read data into R
df <- read.csv("2-data.csv", stringsAsFactors = FALSE)

## preview first six rows of data
head(df)
```

The data set `df` contains seven variables:

- `age`: Age of the respondent
- `race`: Race of the respondent
- `sex`: Sex of the respondent
- `polar`: Affective political polarization (favorability of in-group `-` favoriability of out-group where the groups are D and R political parties)
- `sm_use`: Social media use (extent to which respondent uses social media)
- `network_agree`: Network agreement (extent to which social media networks are politically like-minded)
- `pp`: Political participation (number of political activities engaged in the past 6 months)

## Instructions
Respond to the following three prompts by entering the appropriate code or text below each.

## Prompts

### 1) What is the breakdown (frequencies) of the respondents by `race` and by `sex`? (Hint: create frequency tables `table()` for the `df$race` and `df$sex` variables.)

```{r}
## your code goes here

```

### 2) What is the **mininum** value, **maximum** value, **mean**, and **standard deviation** for the `age`, `polar`, `sm_use`, `network_agree`, and `pp` variables?

```{r}
## your code goes here

## use comment lines like this one to annotate (explain) your code

```

### 3) Write a description of the `age`, `sex`, and `race` variables like you would expect to see in an academic manuscript. It should be approximately one paragraph (4-8 sentences). I've written the first sentence for you---though you will need to insert the correct number of participants.

<!-- to respond to this prompt, edit and add to the text below --> 

An online questionnaire was distributed to undergraduates (N = `__`) enrolled in a general education course at a major midwestern university.


<!-- when you're finished responding to the prompts, make sure to click 'Knit` one last time. then upload the rendered html file (saved every time you `knit` the document) to Canvas -->
