---
title: "Homework #2"
subtitle: "JOURN 8006"
author: "\\<enter your name here>"
date: "3/06/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

For this homework assignment, we will use the same data as the prevoius homwork assignment. So, create a new folder and save the `3-correlation.Rmd` and `2-data.csv` files in it.

## Data

Read in the data file (`2-data.csv`) using the code below.

```{r}
## read data into R
df <- read.csv("2-data.csv", stringsAsFactors = FALSE)

## preview first six rows of data
head(df)
```

The data set `df` contains seven variables:

- `age`: Age of the respondent
- `race`: Race of the respondent
- `sex`: Sex of the respondent
- `polar`: Affective political polarization (favorability of in-group `-` favoriability of out-group where the groups are D and R political parties)
- `sm_use`: Social media use (extent to which respondent uses social media)
- `network_agree`: Network agreement (extent to which social media networks are politically like-minded)
- `pp`: Political participation (number of political activities engaged in the past 6 months)

## Instructions
Respond to the following three prompts by entering the appropriate code or text below each.

## Prompts

### What's the correlation between `sm_use` and `network_agree`

```{r}
## your code goes here

```

### Plot the correlation between `sm_use` and `network_agree`

```{r}
## your code goes here

```


### What is the correlation between `polar` and `pp`

```{r}
## your code goes here

```


### Plot the correlation between `polar` and `pp`

```{r}
## your code goes here

```


### Using the code below, answer the following three questions about the correlation between the weight and mpg of cars (using the mtcars data set).
#### a) What is your guess for the `r` value?
#### b) Is the relationship positive or negative?
#### c) Is the relationship strong or weak?

```{r}
## load ggplot2
library(ggplot2)

## scatter plot of mpg by wt
ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point()
```



