---
title: "Homework #4"
subtitle: "JOURN 8006"
author: "\\<enter your name here>"
date: "4/06/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

For this homework assignment, we will a new data set. So, create a new folder and save the `4-lm.Rmd` and `4-data.csv` files in it.

## Data

Read in the data file (`2-data.csv`) using the code below.

```{r}
## read data into R
df <- read.csv("4-data.csv", stringsAsFactors = FALSE)

## preview first six rows of data
head(df)
```

The data set `df` contains six variables:

- `total_bill`: Total cost of restaurant bill
- `tip`: Amount of tip
- `sex`: Sex of customer (Female, Male)
- `smoker`: Whether customer is a smoker
- `day`: Day of the week (Fri, Sat, Sun, Thur)
- `time`: Meal of the day (Dinner, Lunch)
- `size`: Number of people in party

## Instructions
Respond to the following three prompts by entering the appropriate code or text below each.

## Prompts

### MODEL 1

#### a. Write the formula for a model predicting `tips` (outcome) by `total_bill` (predictor)

```{r}
## your code goes inside the parentheses in the line below
f1 <- formula()
```

#### b. Estimate a model using formula `f1` and data `df`. Store the output as `m1`.

```{r}
## your code goes inside the parentheses in the line below
m1 <- lm()
```

#### c. Summarize `m1`

```{r}
## your code goes here

```

### MODEL TWO

#### a. Write the formula for a model predicting `tips` (outcome) by `total_bill`, `sex`, `smoker`, `day`, `time`, `size` (predictors).

```{r}
## your code goes inside the parentheses in the line below
f2 <- formula()
```

#### b. Estimate a model using formula `f2` and data `df`. Store the output as `m2`.

```{r}
## your code goes inside the parentheses in the line below
m2 <- lm()
```


#### c. Summarize `m2`

```{r}
## your code goes here

```

#### d. Which variables were significant predictors of `tips`? And were the significant predictors positive or negative?

<!-- write your answer here -->



