---
title: "Week 10"
subtitle: "Correlations"
author: "Michael W. Kearney📊<br/>School of Journalism <br/>Informatics Institute <br/>University of Missouri"
date: "<table style=\"border-style:none;padding-top:30px;\" class=\".table\">
  <tr>
    <th style=\"padding-right:75px!important\"><a href=\"https://twitter.com/kearneymw\"> <i class=\"fa fa-twitter\"></i> </a></th>
    <th style=\"padding-left:75px!important\"><a href=\"https://github.com/mkearney\"> <i class=\"fa fa-github\"></i> </a></th>
  </tr>
  <tr style=\"background-color:#fff\">
    <th style=\"padding-right:75px!important\"><a href=\"https://github.com/mkearney\"> @kearneymw </a></th>
    <th style=\"padding-left:75px!important\"><a href=\"https://github.com/mkearney\"> @mkearney </a></th>
  </tr></table>"
output:
  xaringan::moon_reader:
    css: ['robot', 'robot-fonts']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, collapse = TRUE, comment = ">")
options(htmltools.dir.version = FALSE)
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
head <- function(x, n = 6) {
  x <- utils::head(x, n)
  knitr::kable(x, format = "html")
}
```

# \#rstats

Visualizations and data wrangling done using the [tidyverse](https://tidyverse.org).

+ **{ggplot2}**: data visualizations
+ **{dplyr}**: data wrangling

```{r}
## load the tidyverse
library(tidyverse)
```

---

# Correlation

A **correlation** describes how variation in one variable relates to variation in another variable. 

In statistics, correlations describe the **magnitude** and **direction** of a relationship between two variables. 


---

```{r}
## 20 observations from random normal distribution
x <- rnorm(20)
## x + 20 observations from random normal distribution
y <- rnorm(20) + x
## scatter plot
plot(x, y)
```

---

# Correlation coefficient

In statistics, a **coefficient** refers to a numeric estimate.

A **correlation coefficient** `r` is a number that describes a correlation.

There are several types of correlation coefficients but they all tend to assume the following:

+ The **range of possible values** goes from -1 to +1
+ **Direction** is described by the sign of the coefficient (negative = negative relationship and vice versa)
+ **Magnitude** is described by the coefficient's distance from zero
    + A correlation of zero indicates an absence of correlation (no relationship)

---

# Types of correlations

**Pearson product-moment**: linear relationship between two variables 
$$x\sim\sim y$$

**Spearman's rho**: relationship between rankings of two variables 
$$x\sim\sim y$$

**Intraclass**: relationship between paired observations 
$$x_{t1}\sim\sim x_{t2}$$

---

# Semester

All of the correlations are similar, but you can assume when I talk about correlations I'm talking about the Pearson product-moment correlation.

When your variables are ordinal (ordered categories), it's important to remember there's a better way to estimate the correlation coefficient.

---

# Interval/ratio data

```{r}
## 20 observations from random normal distribution
x <- rnorm(20)
## x + 20 observations from random normal distribution
y <- rnorm(20) + x
## create data frame
df <- data.frame(x, y)

## print
head(df)
```



---

# Ordinal data

```{r}
## sample twenty values from numbers 1,2,3,4,5,6
x <- sample(c(1, 2, 3, 4, 5, 6), 20, replace = TRUE)
## sample twenty more, add them with x, and then divide by 2
y <- round((x + sample(c(1, 2, 3, 4, 5, 6), 20, replace = TRUE)) / 2, 0)
## create data frame
df_ordered <- data.frame(x, y)

## print
head(df_ordered)
```


---

# `cor(method = "pearson")`

By default, `cor()` returns the Pearson product-moment correlation.

```{r}
## using cor defaults
cor(df$x, df$y)
```

Use `method` to specify the type of correlation and `use` to deal with missing data

```{r}
## pearson product-moment correlation
cor(df$x, df$y, method = "pearson", use = "complete.obs")
```

---

# `cor(method = "spearman")`

```{r}
## pearson product-moment correlation
cor(df_ordered$x, df$y, method = "pearson", use = "complete.obs")

## spearman rho correlation
cor(df_ordered$x, df$y, method = "spearman", use = "complete.obs")
```


---

# Correlation tool

To visualize correlations of different values (-1 to 1) check out this Shiny app I made: 

+ https://mikewk.shinyapps.io/correlation/

`---
# Practice

Guess the correlation coefficient (the game):

+ http://guessthecorrelation.com/

