---
title: "Week 2"
subtitle: "Variables"
date: "`r Sys.Date()`"
author: "Michael W. Kearney, PhD"
institute: |
  | School of Journalism
  | Informatics Institute
  | University of Missouri
output: uslides::uslides_document
---

```{r setup, include=FALSE}
knitr::knit_hooks$set(mysize = function(before, options, envir) {
  if (before) return(options$size)
})
knitr::opts_chunk$set(collapse = TRUE, mysize = TRUE, size = "\\scriptsize")
knit_print.character = function(x, ...) {
  sQuote(noquote(x))
}
```

# Recap

## Scientific method

- Try to be "objective"
- Theories should be [falsifiable](http://mikewk.com)
- Research should be reproducible
- Knowledge is cumulative and provisional

# Variables

## Variable

- A **constant** is a fixed value that never changes.
    - e.g., pi, the number 1, etc.
- A **variable** is a value that differs across observations.
    - can often be thought of as features or characteristics

## Variable values

- **Values** are measurements (observations) on a given variable
    - e.g., Tracy's height (variable) is `6'6''` (value)
    - e.g., Avery's final race position (variable) is `1` (value)
    - e.g., Cory's skill level in chess (variable) is `master` (value)
    - e.g., Rory's hometown (variable) is `Kansas City` (value)
- Different levels of measurement enable different levels of analysis

## Levels of measurement
- Nominal
- Ordinal
- Interval
- Ratio

## Ratio/numeric/interval

Numeric variables are [or represent] real numbers.

```{r}
foo <- function(x) {
  rnorm(x)
}
foo(5)
class(c(1.25, 3.5, 4))
```

> Technically, if there's a true zero, then it's considered "ratio". Otherwise, it's "interval".

## Ordinal/integer

Ordinal variables are a meaningful sequence of integers.

```{r}
class(1:3)
```

## Categorical/character

Categorical variables are used to represent nominal-level categories.

```{r}
class(c('a', 'b', 'c'))
```


## Defining variables

**Conceptual definition**

- A description of a variable's **theoretical** meaning.

**Operational definition**

- A description of a variable's **observable** meaning.

# Example

## Size

Define the word size

- **Size** refers to the dimensions of an object

## Class 1

## Class 2
