---
title: "Week 2"
subtitle: "Variables"
date: "`r Sys.Date()`"
author: "Michael W. Kearney, PhD"
institute: |
  | School of Journalism
  | Informatics Institute
  | University of Missouri
output: uslides::uslides_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, size = "\\scriptsize")
```

# Recap

## Scientific method

- Try to be "objective"
- Theories should be [falsifiable](http://mikewk.com)
- Research should be reproducible
- Knowledge is cumulative and provisional

# Sampling

## Random sample

- Technically, it should be a "uniform" random sample
- Everyone in a sampling frame (all possible data points) has an equal probability of being selected
- Best method for making inferences but expensive and difficult to do

## Quasi-random sample

- Samples that use mathematical rules but lack access to the full population
- One common technique is **probability matching**, which is when you match the demographics in a sample with the desired population

## Snowball sample

- When you use one to recruit more into the sample
- Not very representative, but useful for getting access to niche or hard-to-reach groups

## Convenience sample

- A sample that is not random but selected largely due to convenience (low cost, ease of access, etc.)

## Why care about sampling?

- The sampling method and study design has a direct effect on one's ability to make **inferences** from data
- **Inferences** are conclusions drawn from a sample and applied to a population

# Variables

## Variable

- A **constant** is a fixed value that never changes.
    - e.g., pi, the number 1, etc.
- A **variable** is a value that differs across observations.
    - can often be thought of as features or characteristics

## Variable values

- **Values** are measurements (observations) on a given variable
    - e.g., Tracy's height (variable) is `6'6''` (value)
    - e.g., Avery's final race position (variable) is `1` (value)
    - e.g., Cory's skill level in chess (variable) is `master` (value)
    - e.g., Rory's hometown (variable) is `Kansas City` (value)
- Different levels of measurement enable different levels of analysis

## Levels of measurement
- **Nominal**: values represent different categories [or named things]
- **Ordinal**: values represent meaningful sequence
- **Interval**: values represent meaningful sequence using equi-distant intervals
- **Ratio**: values represent real numbers

## Nominal
Values represent different categories [or named things]

- Can be used to operationalize anything
- This is often done using *dummy codes*


## Ordinal
Values represent meaningful sequence

- The order people finish in a race
- The distance from 1st to 2nd can vary wildly with the distance from 2nd to 3rd

## Interval
Values represent meaningful sequence using equi-distant intervals

- Likert-type items, e.g., *I always make my bed in the morning: Strongly Agree... Strongly Disagree* and other survey items that measure a range of feelings/attitudes using numbers
- This is why the visual representation of numbers on a survey is often important


## Ratio 
Values represent real numbers

- Numbers correspond to some non-arbitrary meaning.
- A true 0 (zero) exists

## Temperature
**Do the statements below add up?**

- The temperature today is 30 degrees Fahrenheit
- The temperature yesterday is 60 degrees Fahrenheit
- Yesterday was twice as hot as tomorrow

## Numeric

- In R, we will refer to **continuous** interval and ratio variables as `numeric` values.
- Values are continuous if they don't only exist as discrete units

```{r}
x <- c(1.25, 3.5, 4)
class(x)
```

## Integer

- In R, we will refer to **discrete** numbers as `integer` values.

```{r}
x <- c(1, 2, 3)
class(x)
```

## Categorical/character

- I will tend to refer to nominal variables as categorical variables
- Categorical variables are represented in R as `character` and `factor` variables

```{r}
x <- c('a', 'b', 'c')
class(x)

x <- as.factor(x)
class(x)
```


## Defining variables

**Conceptual definition**

- A description of a variable's **theoretical** meaning.

**Operational definition**

- A description of a variable's **observable** meaning.

# Example

## Size

Define the word size

- **Size** refers to the dimensions of an object

## Two classes

Let's say we want to find the size of two classes

- **CLASS 1**: 20x20 room with 5 students
- **CLASS 2**: 10x10 room with 10 students 

## Defining "class size"

- **CLASS 1**: 20x20 room with 5 students
- **CLASS 2**: 10x10 room with 10 students 
- If we define the size of a class in terms of physical space, class 1 is bigger
- If we define the size of a class in terms of number of students, class 2 is bigger
- It's even possible to define class in terms of mass of students

Is any definition right or wrong?


## Practice

- Let's practice conceptually and operationally defining variables


## Media exposure

- Conceptual definition:
- Operational definition:

## Views on climate change

- Conceptual definition:
- Operational definition:

## Coverage of environmental issues

- Conceptual definition:
- Operational definition:

## Accuracy in reporting

- Conceptual definition:
- Operational definition:

## Fake news

- Conceptual definition:
- Operational definition:

## Social media use

- Conceptual definition:
- Operational definition:

## Political ideology

- Conceptual definition:
- Operational definition:

## Perceived role of journalists

- Conceptual definition:
- Operational definition:

# Independent and dependent variables

## Classifying variables

- Variables can be classified according to their status in a theoretical and/or statistical model
- There are several different labels applied toward this end, but for the most part they all do the same thing
- Variables that represent the *cause*
- Variables that represent the *effect*

## Terminology

- The cause is the **independent** variable
- The effect is the **dependent** variable
- There's also exogeneous and endogenous, predictors and outcomes, and probably some other synonyms

## Cause isn't always cause

- Although we use the cause/effect heuristic when modeling the relationship between two variables, we usually acknowledge multiple and/or reinforcing causes or some underlying "root cause"
- As long as we all know that, the cause/effect heuristic is still helpful for organizing our theoretical and statistical models

## Practice

- Let's practice identifying independent and dependent variables

## Independent or dependent?

- Sun
- Surface temperature

## Independent or dependent?

- Reading
- Knowledge

## Independent or dependent?

- Public perceptions of an event
- Reporting information about an event

## Independent or dependent?

- Breakfast
- Health

## Independent or dependent?

- Exercise
- Health

## Independent or dependent?

- Social media
- Face-to-face interaction


# Hypotheses/research questions

## Relationships between variables

- Research questions ask about the relationship between two or more variables
    - These questions can guess direction (e.g., positive vs negative) but that kind of defeats the purpose
- Hypotheses predict the relationship between two or more variables
    - Hypotheses can either predict **whether** there is a relationship or what **direction** the relationship is in

## RQ/H

- Use research questions when there isn't a lot of research in an area and the theory isn't clear about what to expect
- Otherwise, use hypotheses

# Research designs

## Overview of study designs

- Surveys
- Experiments
- Content/text analysis

## Next week

- Survey research with Dr. Ben Warner from Communication Department
- Make sure you've completed your IRB training
