---
title: "Week 7"
subtitle: "Descriptives"
date: "`r Sys.Date()`"
author: "Michael W. Kearney, PhD"
institute: |
  | School of Journalism
  | Informatics Institute
  | University of Missouri
output: 
  uslides::uslides_html:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
Sys.setenv(
  USLIDES_BOTTOMLEFT = "uslidelogos/bl.jpg",
  USLIDES_BOTTOMRIGHT = "uslidelogos/br.jpg",
  USLIDES_BANNER = "uslidelogos/bn.jpg",
  USLIDES_BACKGROUND = "uslidelogos/bg.jpg"
)
```

<!-- ############################################# -->
<!-- ##             RECAP/OVERVIEW              ## -->
<!-- ############################################# -->

# Recap and overview

## #rstats
> + **Software**
>     - **R**: A statistical programming language
>     - **Rstudio**: graphical interface (development environment)
> + **R objects and classes**
>     - Creating and overriding R objects
>     - Classes: `numeric`, `integer`, `character`, `factor`, `data.frame`
> + **R functions**
>     - Named automated tasks used on data (e.g., `mean(x)`)

## Due today
+ Upload output from Homework #1

## Overview
+ **Descriptives**: summarizing data
    - Measures of central tendency
    - Measures of dispersions
+ Getting data *into* R
    - Reading different types of data files
+ Descriptives in **R**
    - Summary functions/packages and output


<!-- ############################################# -->
<!-- ##             DESCRIPTIVES                ## -->
<!-- ############################################# -->

# Descriptives

## "Statistics"
+ **Descriptive** statistics (AKA *parametrics*)
    - Describing the population of data
    - Not a lot of probability theory required
+ **Inferential** statistics (AKA *statistics*)
    - Describing the population using a sample of data
    - Probability theory is key

## Descriptive statistics
+ Measures of **central tendency**
    - Describes the **middle** of the data
+ Measures of dispersion
    - Describes the **spread** of the data

## Central tendency
+ **Mean**: the expected value (often called "average")
+ **Median**: the mid-point of the data
+ **Mode**: the most common data point

## Mean
+ Calculate the mean `mean()` by summing `sum()` all values and dividing `/` by the number of values `length()`

```{r}
## 10 numbers from random distribution with mean of 0
x <- rnorm(10, mean = 0)

## calculate mean
sum(x) / length(x)

## use mean function
mean(x)
```


## Median
+ Calculate the median `median()` by arranging `sort()` all values and find the middle (or values *tied* for the middle) observation.

```{r}
## sort the numbers by magnitude
x <- sort(x)

## print and find the middle values (5th and 6th)
x

## use median function
median(x)
```


## Mode
+ Calculate the mode by creating a frequency table `table()` and sorting `sort()` in descending order (biggest to smallest)

```{r}
## sample 100 values from the series of numbers 1:10 (1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
y <- sample(1:10, 100, replace = TRUE)

## create frequency table
y_freq <- table(y)

## sort and print table
sort(y_freq, decreasing = TRUE)
```


## Important R note
+ In R missing values are coded as `NA`
+ Inputs containing `NA` will return `NA` for these functions:
    - `sum()`, `mean()`, `median()`, `min()`, `max()`, `range()`, `var()`, `sd()`
+ To avoid this, include the argument `na.rm = TRUE`
+ Although not always, this (ignoring missing values) is usually what you want to happen

## `na.rm = TRUE`
+ Example of finding mean when `x` contains a missing value (`NA`)
```{r}
## add missing value ot x
x <- c(NA, x)

## returns NA
mean(x)

## returns the mean we want
mean(x, na.rm = TRUE)
```

## Dispersion
+ **Range**: the minimum and maximum values. Often expressed as a distance between the two.
+ **Variance**: distance from the mean
+ **Standard deviation**: distance from the mean expressed in standardized units

## Range
+ Calculate the range `range()` by finding the maximum `max()` and minimum `min()` values.

```{r}
## min and max values of y
min(y)
max(y)

## calculate distance between the two
max(y) - min(y)

## use the range function
range(y)
```


## Variance
+ Calculate the variance `var()` by summing `sum()` the squared `^2` distance `-` from the mean `mean()` and dividing `/` by the number of observations `length(x) - 1`

```{r}
## calculate variance
sum((y - mean(y))^2) / (length(y) - 1)

## use the var function
var(y)
```

## Standard deviation
+ Calculate the standard deviation `sd()` by taking the square root `sqrt()` of the variance `var()`.

```{r}
## calculate variance
yvar <- var(y)

## square root of variance
sqrt(yvar)

## use sd function
sd(y)
```

<!-- ############################################# -->
<!-- ##         GETTING DATA INTO R             ## -->
<!-- ############################################# -->

# Getting data into R

## CSV
+ **CSV**: comma separated value
```{r}
read.csv("data/csv.csv")
```

## \*SV/delimited
```{r}
## uncomment following line for help documentation
#?read.table
read.delim("data/tsv.tsv")
```

## dat
+ Wikipedia describes `.dat` as
> not [a] specific file type, often generic extension for "data" files for a variety of applications
+ Often associated with Mplus (statistical software program)
+ See: \*SV methods, i.e., `read.table()`

## xlsx (Excel)
```{r}
## install readxl package if not already
if (!requireNamespace("readxl", quietly = TRUE)) {
  install.packages("readxl")
}
## read an excel file
readxl::read_excel("data/xlsx.xlsx")
```

## .sav (SPSS)
```{r}
## install readxl package if not already
if (!requireNamespace("haven", quietly = TRUE)) {
  install.packages("haven")
}
## read spss (sav) file
haven::read_spss("data/sav.sav")
```

## dta (Stata)
```{r}
## install readxl package if not already
if (!requireNamespace("haven", quietly = TRUE)) {
  install.packages("haven")
}
## read stata (dta) file
haven::read_stata("data/dta.dta")
```

## rds (R)
```{r}
readRDS("data/rds.rds")
```

<!-- ############################################# -->
<!-- ##           DESCRIPTIVES IN R             ## -->
<!-- ############################################# -->

# Write up

## Data
+ After data collection and prior to analysis, research projects should describe, or summarize, the data
+ This should be included in the **methods** section of your paper
+ Descriptives for:
    - Participants
    - Variables

## Participants
+ Descriptives in the data summary typically include the following
    - Number of observations (total number `N = 345`)
    - Demographic (age, sex, race, education, etc.) breakdowns (numbers and percents *15.3% were female* (`n = 23`))
        - For age, usually a range and the mean

## Variables
+ Study variable information: 
    - min
    - max
    - mean
    - sd
+ Correlation table (don't worry about that yet)

## Useful functions/packages
+ Fortunately, there are lots of good functions and packages to choose from
    - `summary()` (base R)
    - `describe()` from **{psych}**
    - `skim()` from **{skimr}**
    - `summarize()` from **{rockchalk}**


```{r}
```


